--- 
title: "Data FAIRy"
author: "Anton Van de Putte, Yi-Ming Gan, Maxime Sweetlove"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
description: "Biological data publishing guidelines"
---

# About {-}

This document serves to guide researchers who wish to manage the biological data from their scientific expedition

Issues or suggestions about the content can be raised through the [issue tracker](https://github.com/79-6d/data-fairy/issues) (GitHub account required).

## How to build this book locally {-}

This material has been written using [bookdown](https://github.com/rstudio/bookdown) and [R](https://cran.r-project.org/). The dependencies of this project is managed using [renv](https://rstudio.github.io/renv/).

To build the book locally, clone the [repo](https://github.com/79-6d/data-fairy.git).

If dependencies are not automatically installed by `renv` when you open `data-fairy.Rproj`, try the following command.

```{r eval=FALSE}
renv::restore()
```

then run the following lines to build the book:

```{r echo=FALSE}
library(here)  # placed in separate chunk so that renv picks up the pkg
```

```{r eval=FALSE}
setwd(here())  # set working directory to project root directory 
bookdown::render_book("index.Rmd")
```

And view it with:

```{r}
browseURL("docs/index.html")
```

